HyperLogLog: Exponentially Bad in Adversarial Settings

### 2020-05-05 docs/HLL_exp_bad/main.tex
draft of S1

\begin{itemize}
\item[S1:] The adversary has access to a \textit{test oracle} provided by the sketch owner, which allows it to submit an item and learn the hypothetical effect of inserting it into the sketch. It is then able to determine if inserting this item \textit{would} increase the cardinality estimate, without actually inserting it.
%The adversary is able to test the hypothetical effect of the insertion of an item on the cardinality estimation of the sketch. This can achieved via a \textit{test oracle} provided by the sketch owner, which allows an user to test the effect of an insertion on the cardinality estimate but without actually inserting the item, or if the attacker would gain access to a black-box replica of the sketch in use, i.e., buying the same switch and testing in the lab. %black-box access to the HLL in use -- that is, the adversary can ask for the HLL's cardinality estimate at any point in time, and so is able to check the effect on the cardinality estimate when inserting an item.
%\kenny{TODO: clarify more what this is terms of a ''test oracle''; its stronger than described because the adversary can test the effect of an insertion without making the insertion!s}
\item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''. In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack. This kind of attack scenario is appropriate in the context of the port scanning attack of~\cite{portscanhll}.
\item[S3:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use. Additionally, it can access and interact with the sketch via an API provided by the sketch owner, allowing it to insert items into the HLL sketch and ask for the HLL's cardinality estimate at any point in time.
\item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch. It then causes chosen inputs to be inserted into the HLL sketch. An S4 adversary can compute a lower bound on the HLL cardinality estimate based on its snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
\end{itemize}

@@

We start by highlighting the fact that S1 seems unrealistic in practice, as there is no reason why a typical HLL implementation would provide such a test oracle. The authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker gains access to a replica of the machine on which the HLL is implemented. The adversary could run tests on it, sequentially inserting, asking for the cardinality estimation and resetting the device, retaining only the items which do not increment the HLL estimate. This may be possible under certain circumstances, but we do not consider it typical of HLL deployments.

>>>>

\begin{itemize}
\item[S1:] The adversary has access to an empty black-box replica of the sketch in use. He can use the API to insert elements into that shadow device, get the cardinality estimate, and reset the sketch. The attacker does not know the elements previously added by honest users nor the state. That is we can test and find elements that are unlikely to increase the HLL estimate on the shadow HLL and then insert the on the target HLL but we do not know the state of the target HLL. It would also be reasonable for the provider to transform the items/hash with  a salt e.g. with a different salt per customer. So one can have both an open source implementation and not know the hash.
\item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''. In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack. This kind of attack scenario is appropriate in the context of the port scanning attack of~\cite{portscanhll}.
\item[S3:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use. Additionally, it can access and interact with the sketch via an API provided by the sketch owner, allowing it to insert items into the HLL sketch and ask for the HLL's cardinality estimate at any point in time.
\item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch. It then causes chosen inputs to be inserted into the HLL sketch. An S4 adversary can compute a lower bound on the HLL cardinality estimate based on its snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
\end{itemize}

@@

We start by highlighting the fact that S1 seems unrealistic in practice. In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker learns which machine is used to do the monitoring and buys the same device. It is reasonable to think that the provider may use a different salt per device, which would make any effect on the shadow device unexpected on the target and would break the whole attack presented in Section 3.2 of \cite{hllvuln}. We consider that the machines use the same exact internals. Assuming that such setup is possible, we show that it is actually possible for the attacker to learn the internals of the HLL sketch. The API must explicitly tell the customer the error of the sketch, and hence the number of buckets. We show that the attacker can learn the hash of a targeted item using the knowlwdege of S1, and thus end up in the S2 scenario. To learn $h(x)$, the attacker can insert the item $x$ on en empty sketch, and ask for the resulting cardinality, which we denote $c$. By solving:
\begin{eqnarray*}
    \alpha 2^{2n}(\sum_{j=1}^{2^n}2^{-c_j})^{-1}& = & c \\
    \Leftrightarrow log_2(c)-log_2(\alpha m^2 - cm + c) &=& M[i].
\end{eqnarray*}
where  the attacker can learn the value of the updated estimator of the bucket $x$ was mapped to, and recover some bits of $h(x)$. This may be possible under certain circumstances, but we do not consider it typical of HLL deployments, and furthermore, it is easy for the attacker to escalate to the knowledge of S2.
